---
title: "Files"
author: "Rory Nolan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

First let's load the library: 
```{r load}
library(filesstrings)
```

## Move files around
I find it bizarre that base R has no `file.move`. To move a file, you have to unintuitively rename it. `filesstrings` provides `file.move(files, destinations)`. This function has the nice feature that if you try to move files to a directory that doesn't exist, it creates the directory first and then puts the files inside. 
Let's create a directory and a file:
```{r create dir}
dir.create("tmp_dir")
file.create("tmp.txt")
```
Now let's put the file into the directory:
```{r put a file in a dir}
file.move("tmp.txt", "tmp_dir")
```

## Delete Directories
To delete directories with base R, one has to use `unlink(..., recursive = TRUE)`. The `filesstrings` package gives you `dir.remove()` which does the same job.
```{r, remove directory}
dir.remove("tmp_dir")
```

## Remove spaces from file names
"A space in your file name is a hole in your soul." - Jenny Bryan

`RemoveFileNameSpaces(replace.with = "_")` replaces them all with underscores for all files in a directory. By default, they are replaced with nothing.
```{r, RemoveFileNameSpaces}
file.create(c("file 1.txt", "file 2.txt"))
RemoveFileNameSpaces(pattern = "txt$", replace.with = "_")
list.files(pattern = "txt$")
file.remove(list.files(pattern = "txt$"))  # clean up
```

## Messed up file numbering
The microscope I use numbers files with 3 numbers by default, i.e. `file001.tif`, `file002.tif` and so on. This is a problem when the automatic numbering passes 1000, whereby we have `file999.tif`, `file1000.tif`. What's the problem with this? Well, sometimes you need alphabetical order to reflect the true order of your files. These file numbers don't satisfy this requirement:
```{r NiceNums setup}
file.names <- c("file999.tif", "file1000.tif")
sort(file.names)
```
so `file1000.tif` comes before `file999.tif` in alphabetical order. The function `NiceNums` returns the names that we'd like them to have:
```{r NiceNums}
NiceNums(file.names)
```
The function `NiceFileNums` applies such renaming to all the files in an entire directory. It wraps `NiceNums`.

## The name of a file without the extension
```{r BeforeLastDot}
BeforeLastDot("spreadsheet_92.csv")
```

## Ensure that a file name has a given extension
Add a file extension if needed:
```{r add file extension 1}
GiveExt("xyz", "csv")
```
If the file name has the correct extension already, it's left alone:
```{r add file extension 2}
GiveExt("xyz.csv", "csv")  
```
Change a file extension:
```{r change file extension}
GiveExt("abc.csv", "txt")  # tack the new extension onto the end
GiveExt("abc.csv", "txt", replace = TRUE)  # replace the current extension
```
